Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}

### 2008-04-17 docs/Godunov_notes/paper.tex
A few minor edits John suggested, and some of my own.

These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
\section{Differences Between These Notes and the Code}
\begin{enumerate}
\item Concerning both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), in the code we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$.
\end{enumerate}
\section{Questions}
\begin{enumerate}
\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity to r-faces.  For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
\end{enumerate}
\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
\begin{enumerate}
\item {\tt VARDEN} does not use the concept of $w_0$, so there is no such thing as $\Ubt$ either.  {\tt VARDEN} only does Cartesian problems.
\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$.
\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, but not $\Ub^{\edge}$
\item The transverse derivative terms in {\tt VARDEN} are done in a conservative formulation, rather than advective.
\end{enumerate}
\cleardoublepage
\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.  These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.  We will only compute the normal component of velocity at each face.\\

@@

The force terms are:
\begin{equation}
\fb =
\left(\begin{array}{c}
f_u \\
f_w \\
\end{array}\right)
= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
\end{equation}
\subsection{2D Cartesian Case}
We begin by computing a time-centered 1D extrapolation of $u$ to r-faces:
\begin{eqnarray}
u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
\end{eqnarray}
Then we upwind based on $w^{\trans}$:
\begin{equation}
u_{\ib-\half\eb_r}^{1D} =
\begin{cases}
\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
\end{cases}
\end{equation}
Then, we do a full-dimensional extrapolation of $u$ to x-faces:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{\mac} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
u_{R,\ib-\half\eb_x}^{\mac} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
\end{eqnarray}
Finally, we solve a Riemann problem:
\begin{equation}
u_{\ib-\half\eb_x}^{\mac,*} =
\begin{cases}
0, & \left(u_{L,\ib-\half\eb_x}^{\mac} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac}\right| < \epsilon, \\
u_{L,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} > 0, \\
u_{R,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} < 0. 
\end{cases}
\end{equation}

@@

\wt_{L,\ib-\half\eb_r}^{\mac} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\

@@

\wt_{R,\ib-\half\eb_r}^{\mac} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\

@@

0, & \left(w_L^{\mac} \le 0 ~ {\rm AND} ~ w_R^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\mac} + w_R^{\mac}\right| < \epsilon, \\
\wt_{L,\ib-\half\eb_r}^{\mac}, & w_L^{\mac} + w_R^{\mac} > 0, \\
\wt_{R,\ib-\half\eb_r}^{\mac}, & w_L^{\mac} + w_R^{\mac} < 0. 

@@

Finally, we call {\tt addw0} to compute $\Ub^{\mac,*}$ from $\Ubt^{\mac,*}$.

@@

If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}$ to $h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.  Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.

@@

\section{Computing $\Ub^{\mac}$ in {\tt VARDEN}}

@@

Now we do a full-dimensional extrapolation to get only the normal velocity at each face:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{\mac} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
&-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
u_{R,\ib-\half\eb_x}^{\mac} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
&-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
\end{eqnarray}
Then, solve a Riemann problem:
\begin{equation}
u_{\ib-\half\eb_x}^{\mac} =
\begin{cases}
0, & \left(u_{L,\ib-\half\eb_x}^{\mac} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac}\right| < \epsilon, \\
u_{L,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} > 0, \\
u_{R,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} < 0.
\end{cases}
\end{equation}
Use the exact analogous procedure to obtain $v^{\mac}_{\ib-\half\eb_y}$.
%\subsection{3D Cartesian Case}
\cleardoublepage
\section{Computing $\Ub^{\edge}$ and $\rho^{\edge}$ in {\tt VARDEN}}
To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.  To the algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.  
\subsection{2D Cartesian Case}
The 1D extrapolation is:
\begin{eqnarray}
\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \\
\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n.
\end{eqnarray}
Then we upwind based on $u^{\mac}$:
\begin{equation}
\rho_{\ib-\half\eb_x}^{1D} =
\begin{cases}
\half\left(\rho_{L,\ib-\half\eb_x}^{1D} + \rho_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
\end{cases}
\end{equation}
Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:
\begin{eqnarray}
\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - \rho_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
&& \\
\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - \rho_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
\end{eqnarray}
Then, we upwind based on $u^{\mac}$.
\begin{equation}
\rho_{\ib-\half\eb_x}^{\edge} =
\begin{cases}
\half\left(\rho_{L,\ib-\half\eb_x}^{\edge} + \rho_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
\end{cases}
\end{equation}
%\subsection{3D Cartesian Case}
\end{document}

>>>>

These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
\section{Differences Between These Notes and the Code}
\begin{enumerate}
\item Concerning both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), in the code we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$.
\end{enumerate}
\section{Questions}
\begin{enumerate}
\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity to r-faces.  For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
\end{enumerate}
\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
\begin{enumerate}
\item {\tt VARDEN} does not use the concept of $w_0$, so there is no such thing as $\Ubt$ either.  {\tt VARDEN} only does Cartesian problems.
\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$.
\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, but not $\Ub^{\edge}$
\item The transverse derivative terms for $\rho$ in {\tt VARDEN} are done in a conservative formulation, rather than advective.
\end{enumerate}
\cleardoublepage
\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.  These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.  We will only compute the normal component of velocity at each face.\\

@@

The force terms are:
\begin{equation}
\fb =
\left(\begin{array}{c}
f_u \\
f_w \\
\end{array}\right)
= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
\end{equation}
\subsection{2D Cartesian Case}
We begin by computing a time-centered 1D extrapolation of $u$ to r-faces:
\begin{eqnarray}
u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
\end{eqnarray}
Then we upwind based on $w^{\trans}$:
\begin{equation}
u_{\ib-\half\eb_r}^{1D} =
\begin{cases}
\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
\end{cases}
\end{equation}
Then, we do a full-dimensional extrapolation of $u$ to x-faces:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
\end{eqnarray}
Finally, we solve a Riemann problem:
\begin{equation}
u_{\ib-\half\eb_x}^{\mac,*} =
\begin{cases}
0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0. 
\end{cases}
\end{equation}

@@

\wt_{L,\ib-\half\eb_r}^{\mac,*} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\

@@

\wt_{R,\ib-\half\eb_r}^{\mac,*} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\

@@

0, & \left(w_L^{\mac,*} \le 0 ~ {\rm AND} ~ w_R^{\mac,*} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\mac,*} + w_R^{\mac,*}\right| < \epsilon, \\
\wt_{L,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} > 0, \\
\wt_{R,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} < 0. 

@@

We then project $\Ub^{\mac,*}$ to obtain $\Ub^{\mac}$ and we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.

@@

If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}\rightarrow h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.  Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.

@@

\section{Computing $\Ub^{\mac,*}$ in {\tt VARDEN}}

@@

Now we do a full-dimensional extrapolation to get only the normal velocity at each face:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{\mac,*} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
&-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
u_{R,\ib-\half\eb_x}^{\mac,*} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
&-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
\end{eqnarray}
Then, solve a Riemann problem:
\begin{equation}
u_{\ib-\half\eb_x}^{\mac,*} =
\begin{cases}
0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0.
\end{cases}
\end{equation}
Use the exact analogous procedure to obtain $v^{\mac,*}_{\ib-\half\eb_y}$.
%\subsection{3D Cartesian Case}
\cleardoublepage
\section{Computing $\Ub^{\edge}$ and $\rho^{\edge}$ in {\tt VARDEN}}
To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.  To the algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.  
\subsection{2D Cartesian Case}
The 1D extrapolation is:
\begin{eqnarray}
\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \\
\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n.
\end{eqnarray}
Then we upwind based on $u^{\mac}$:
\begin{equation}
\rho_{\ib-\half\eb_x}^{1D} =
\begin{cases}
\half\left(\rho_{L,\ib-\half\eb_x}^{1D} + \rho_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
\end{cases}
\end{equation}
Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:
\begin{eqnarray}
\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - (\rho\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
&& \\
\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - (\rho\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
\end{eqnarray}
Then, we upwind based on $u^{\mac}$.
\begin{equation}
\rho_{\ib-\half\eb_x}^{\edge} =
\begin{cases}
\half\left(\rho_{L,\ib-\half\eb_x}^{\edge} + \rho_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
\end{cases}
\end{equation}
%\subsection{3D Cartesian Case}
\end{document}
